---
output: github_document
bibliography: bib.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
# hose: Higher-Order Spectral Estimators
[![Build Status](https://travis-ci.org/dcgerard/hose.svg?branch=master)](https://travis-ci.org/dcgerard/hose)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/dcgerard/hose?branch=master&svg=true)](https://ci.appveyor.com/project/dcgerard/hose)
[![CRAN status](https://www.r-pkg.org/badges/version/hose)](https://CRAN.R-project.org/package=hose)

# Summary
`hose` is a package designed for working with higher-order spectral
estimators, which were first introduced in @gerard2017adaptive. These
estimators are based on the higher-order singular value decomposition
of @lathauwer2000multilinear and are useful when your data exhibit
tensor-specific structure, such as having approximately low
multilinear rank. This code will allow you to:

- Calculate Stein's unbiased risk estimate (SURE) for all higher-order
  spectral estimators that are weakly differentiable and satisfy mild
  integrability conditions.
- Calculate the mode-specific soft-thresholding estimator that
  minimizes the SURE using a coordinate descent algorithm.
- Iterate through all the possible multilinear ranks of a mean tensor
  and choose the multilinear rank the minimizes the SURE.
- Calculate a generalized SURE, motivated by generalized cross
  validation [@josse2016adaptive], for all higher-order spectral
  estimators.
- Calculate the SURE of estimators that apply higher-order spectral
  shrinkage to sub-tensors of the overall data tensor.
- Calculate the SURE for estimators that individually shrink elements
  of the core array of the HOSVD of the data tensor.
- Non-parametrically estimate the variance to use in these SURE
  procedures.
  
  
The main functions are:

- `get_c()`: Pre-format the data before applying mode-specific singular value 
  shrinkage.
- `tensor_var_est()`: Estimate the variance of the data from multiple options.
- `soft_coord()`: Estimate the underlying low-rank mean tensor via
  soft-thresholding.

# Citation

If you find these methods useful, please cite

> **Gerard, David**, and Peter Hoff. 2017. "Adaptive Higher-Order
> Spectral Estimators." *Electron. J. Statist.* 11 (2). The Institute
> of Mathematical Statistics; the Bernoulli Society:
> 3703--37. <https://doi.org/10.1214/17-EJS1330>.

Or, using BibTex:

``` tex
@ARTICLE{gerard2017adaptive,
    AUTHOR = {David Gerard and Peter Hoff},
     TITLE = {Adaptive higher-order spectral estimators},
   JOURNAL = {Electron. J. Statist.},
  FJOURNAL = {Electronic Journal of Statistics},
      YEAR = {2017},
    VOLUME = {11},
    NUMBER = {2},
     PAGES = {3703-3737},
      ISSN = {1935-7524},
       DOI = {10.1214/17-EJS1330},
      SICI = {1935-7524(2017)11:2<3703:AHOSE>2.0.CO;2-Q},
}
```

# Installation

You can install from CRAN in the usual way:
```{r, eval=FALSE}
install.packages("hose")
```

Or, to install the latest (unstable) version, run the following code in R:
```{r, eval = FALSE}
install.packages(c("tensr", "softImpute", "RMTstat", "devtools"))
devtools::install_github("dcgerard/hose")
```

# Vignette

I've provided a vignette demonstrating the methods available in
`hose`. You can find it
[here](http://dcgerard.github.io/code/sure_example.html). Or you can
build the vignette on install with

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("dcgerard/hose", build_vignettes = TRUE)
```

and access the vignette by running the following code in R:

```{r, eval = FALSE}
utils::vignette("sure_example", package = "hose")
```

# References

